---
title: "Create Stats"
output: html_document
date: "2024-03-05"
purpose: "Calculate Serotype changes in proportion across three vaccine periods stratifying by vaccine and non-vaccine serotype; Calculate GPSC changes in proportion across three vaccine periods stratifying by vaccine and non-vaccine serotype; Calculate AMR changes in proportion across three vaccine periods"
---

# Calculate serotype changes in proportion across three vaccine periods stratifying by vaccine and non-vaccine serotype

## Install and load required packages

These are the packages required for analysis using this R script.

```{r install packages}
#install.packages("plyr")
#install.packages("pwr")

#Load packages
library(plyr)
library(pwr)
library(readr)
```

## Load input data

```{r load dataset}
# set base path for working directory
base_path <- getwd()

# set import file name
merged_data_file_name <- "merged_data.csv"

# set data file path
merged_data_file_path <- file.path(base_path, merged_data_file_name)

# load input data
merged_data <- read_csv(merged_data_file_path)

# rename column Serotype to In_Silico_Serotype for the script
colnames(merged_data)[colnames(merged_data) == "Serotype"] <- "In_Silico_Serotype"

# Subset the data to include the information you need
data <- subset(merged_data, select = c("In_Silico_Serotype", "Vaccine_Period", "Vaccine_Status"))
```

## Create an output dataframe to store the results

```{r Output dataframe}
In_Silico_Serotype_Pre_PostPCV <- data.frame("Vaccine_Period" = 0, "Vaccine_Status" = 0, "In_Silico_Serotype" = 0, "Period1_In_Silico_Serotype" = 0, "Period1_total" = 0, "Period2_In_Silico_Serotype" = 0, "Period2_total" = 0, "p" = 0)
Pre_In_Silico_Serotype <- 0
Pre_other <- 0
Post_In_Silico_Serotype <- 0
Post_other <- 0
Pre_In_Silico_Serotype_proportion <- 0
Pre_other_proportion <- 0
Post_In_Silico_Serotype_proportion <- 0
Post_other_proportion <- 0
```

## Define the Serotypes, vaccine periods and vaccine status (vaccine serotye - PCV and non-vaccine serotype -NVT)

```{r Define Serotypes, PCV and NVTs}
In_Silico_Serotypes <- unique(data$In_Silico_Serotype) # define the serotypes
vaccine_period <- unique(data$Vaccine_Period) # define the vaccine periods
vaccine_status <- unique(data$Vaccine_Status) # define the vaccine status
```

## Subset the data into PCV and NVTs, then into vaccine periods then count and calculate the proportion of each serotypes found among PCVs and NVTs for each vaccine. Perform chi-square test and calculate the power and significance.Adding 0.001 ensures that the denominator is not zero.

```{r Count and Proportions; Power test and Fishers exact test}

# iterate over and subset the data for each vaccine status
for (k in vaccine_status) {
  data_status <- subset(data, data$Vaccine_Status == k)
  # iterate over the vaccine periods data subset into the vaccine statuses
  for (j in vaccine_period) {
    data_status_period <- subset(data_status, data_status$Vaccine_Period != j)
    pcv <- unique(data_status_period$Vaccine_Period)
    for (i in In_Silico_Serotypes) {
      data_status_period_In_Silico_Serotype <- subset(data_status_period, data_status_period$In_Silico_Serotype == i)
      Pre_In_Silico_Serotype <- nrow(subset(data_status_period_In_Silico_Serotype, Vaccine_Period == pcv[1][1]))
      Post_In_Silico_Serotype <- nrow(subset(data_status_period_In_Silico_Serotype, Vaccine_Period == pcv[2][1]))
      Pre_other <- nrow(subset(data_status_period, In_Silico_Serotype != i & Vaccine_Period == pcv[1][1]))
      Post_other <- nrow(subset(data_status_period, In_Silico_Serotype != i & Vaccine_Period == pcv[2][1]))
      
      Pre_In_Silico_Serotype_proportion <- Pre_In_Silico_Serotype / (Pre_In_Silico_Serotype + Pre_other + 0.001) + 0.001
      Post_In_Silico_Serotype_proportion <- Post_In_Silico_Serotype / (Post_In_Silico_Serotype + Post_other + 0.001) + 0.001
      Pre_other_proportion <- Pre_other / (Pre_In_Silico_Serotype + Pre_other + 0.001) + 0.001
      Post_other_proportion <- Post_other / (Post_In_Silico_Serotype + Post_other + 0.001) + 0.001
      
      In_Silico_Serotype <- matrix(c(Post_In_Silico_Serotype, Pre_In_Silico_Serotype, Post_other, Pre_other), nrow = 2, byrow = TRUE) # create matrix to store counts
      In_Silico_Serotype_proportion <- matrix(c(Post_In_Silico_Serotype_proportion, Pre_In_Silico_Serotype_proportion, Post_other_proportion, Pre_other_proportion), nrow = 2, byrow = TRUE) # create matrix to store the proportions
      pwr <- pwr.chisq.test(w = ES.w2(In_Silico_Serotype_proportion), power = 0.8, df = 1, sig.level = 0.05) # calculate the power of the chi-square test with power and significance level
      
      # check if the count is greater than the calculated power
      if (sum(In_Silico_Serotype) > pwr$N) {
        p <- fisher.test(In_Silico_Serotype, alternative)$p.value # Fishers exact test using the counts of serotypes
        pcv_compare <- unique(data_status_period$Vaccine_Period)
        In_Silico_Serotype_Pre_PostPCV <- rbind(In_Silico_Serotype_Pre_PostPCV, c(toString(pcv_compare), k, i, Pre_In_Silico_Serotype, Pre_In_Silico_Serotype + Pre_other, Post_In_Silico_Serotype, Post_In_Silico_Serotype + Post_other, p))
      }
    }
  }
}

In_Silico_Serotype_Pre_PostPCV <- In_Silico_Serotype_Pre_PostPCV[-1, ]
```

## Write the results stratified by vaccine period to a csv file.

```{r Stratifying the results and write the output}
In_Silico_Serotype_VT_Pre_PostPCV <- subset(In_Silico_Serotype_Pre_PostPCV, Vaccine_Status == "PCV")
In_Silico_Serotype_VT_Pre_PostPCV$newcolumn <- p.adjust(In_Silico_Serotype_VT_Pre_PostPCV$p, "BH", n = nrow(In_Silico_Serotype_VT_Pre_PostPCV))
names(In_Silico_Serotype_VT_Pre_PostPCV)[names(In_Silico_Serotype_VT_Pre_PostPCV) == "newcolumn"] <- "adjusted_p"
write.csv(In_Silico_Serotype_VT_Pre_PostPCV, file.path(base_path, "In_Silico_Serotype_VT_Pre_PostPCV.csv"), row.names = FALSE)

In_Silico_Serotype_NVT_Pre_PostPCV <- subset(In_Silico_Serotype_Pre_PostPCV, Vaccine_Status == "NVT")
In_Silico_Serotype_NVT_Pre_PostPCV$newcolumn <- p.adjust(In_Silico_Serotype_NVT_Pre_PostPCV$p, "BH", n = nrow(In_Silico_Serotype_NVT_Pre_PostPCV))
names(In_Silico_Serotype_NVT_Pre_PostPCV)[names(In_Silico_Serotype_NVT_Pre_PostPCV) == "newcolumn"] <- "adjusted_p"
write.csv(In_Silico_Serotype_NVT_Pre_PostPCV, file.path(base_path, "In_Silico_Serotype_NVT_Pre_PostPCV.csv"), row.names = FALSE)
```

------------------------------------------------------------------------

# Calculate GPSC changes in proportion across three vaccine periods stratifying by vaccine and non-vaccine serotype

## Install and load required packages

These are the packages required for analysis using this R script.

```{r install packages}
#install.packages("epiR")
#install.packages("plyr")
#install.packages("pwr")

#Load packages
library(epiR)
library(plyr)
library(pwr)
library(readr)
```

## Load input data

```{r load dataset}
# set base path for working directory
base_path <- getwd()

# set import file name
merged_data_file_name <- "merged_data.csv"

# set data file path
merged_data_file_path <- file.path(base_path, merged_data_file_name)

# load input data
merged_data <- read_csv(merged_data_file_path)

# subset the data to include the information you need
data <- subset(merged_data, select = c("GPSC", "Vaccine_Period", "Vaccine_Status"))
```

## Create an output dataframe to store the results

```{r Output dataframe}
GPSC_Pre_PostPCV <- data.frame("Vaccine_Period" = 0, "Vaccine_Status" = 0, "GPSC" = 0, "Period1_GPSC" = 0, "Period1_total" = 0, "Period2_GPSC" = 0, "Period2_total" = 0, "p" = 0)
Pre_GPSC <- 0
Pre_other <- 0
Post_GPSC <- 0
Post_other <- 0
Pre_GPSC_proportion <- 0
Pre_other_proportion <- 0
Post_GPSC_proportion <- 0
Post_other_proportion <- 0
```

## Define the GPSCs, vaccine periods and vaccine status (vaccine serotype - PCV and non-vaccine serotype -NVT)

```{r Define GPSCs, PCV and NVTs}
GPSCs <- unique(data$GPSC) # define the GPSCs
vaccine_period <- unique(data$Vaccine_Period) # define the vaccine periods
vaccine_status <- unique(data$Vaccine_Status) # define the the vaccine status
```

## Subset the data into PCV and NVTs, then into vaccine periods then count and calculate the proportion of each GPSC found among PCVs and NVTs for each vaccine. Perform chi-square test and calculate the power and significance. Adding 0.001 ensures that the denominator is not zero.

```{r Count and Proportions; Power test and Fishers exact test}
# iterate over and subset the data for each vaccine status
for (k in vaccine_status) {
  # iterate over the vaccine periods data subset into the vaccine statuses
  data_status <- subset(data, data$Vaccine_Status == k)
  for (j in vaccine_period) {
    data_status_period <- subset(data_status, data_status$Vaccine_Period != j)
    pcv <- unique(data_status_period$Vaccine_Period)
    
    # iterate over each serotype subset into the vaccine statuses and vaccine periods; and calculate proportions
    for (i in GPSCs) {
      data_status_period_GPSC <- subset(data_status_period, data_status_period$GPSC == i)
      Pre_GPSC <- nrow(subset(data_status_period_GPSC, Vaccine_Period == pcv[1][1]))
      Post_GPSC <- nrow(subset(data_status_period_GPSC, Vaccine_Period == pcv[2][1]))
      Pre_other <- nrow(subset(data_status_period, GPSC != i & Vaccine_Period == pcv[1][1]))
      Post_other <- nrow(subset(data_status_period, GPSC != i & Vaccine_Period == pcv[2][1]))
      
      Pre_GPSC_proportion <- Pre_GPSC / (Pre_GPSC + Pre_other + 0.001) + 0.001
      Post_GPSC_proportion <- Post_GPSC / (Post_GPSC + Post_other + 0.001) + 0.001
      Pre_other_proportion <- Pre_other / (Pre_GPSC + Pre_other + 0.001) + 0.001
      Post_other_proportion <- Post_other / (Post_GPSC + Post_other + 0.001) + 0.001
      
      GPSC <- matrix(c(Post_GPSC, Pre_GPSC, Post_other, Pre_other), nrow = 2, byrow = TRUE) # create matrix to store counts
      GPSC_proportion <- matrix(c(Post_GPSC_proportion, Pre_GPSC_proportion, Post_other_proportion, Pre_other_proportion), nrow = 2, byrow = TRUE) # create matrix to store the proportions
      pwr <- pwr.chisq.test(w = ES.w2(GPSC_proportion), power = 0.8, df = 1, sig.level = 0.05) # calculate the power of the chi-square test with power and significance level
      
      # Power test; check if the count is greater than the calculated power
      if (sum(GPSC) > pwr$N) {
        p <- fisher.test(GPSC, alternative)$p.value # Fishers exact test using the counts of serotypes
        pcv_compare <- unique(data_status_period$Vaccine_Period)
        GPSC_Pre_PostPCV <- rbind(GPSC_Pre_PostPCV, c(toString(pcv_compare), k, i, Pre_GPSC, Pre_GPSC + Pre_other, Post_GPSC, Post_GPSC + Post_other, p))
      }
    }
  }
}
GPSC_Pre_PostPCV <- GPSC_Pre_PostPCV[-1, ]

```

## Write the results stratified by vaccine period to a csv file.

```{r Stratifying the results and write the output}
GPSC_VT_Pre_PostPCV <- subset(GPSC_Pre_PostPCV, Vaccine_Status == "PCV")
GPSC_VT_Pre_PostPCV$newcolumn <- p.adjust(GPSC_VT_Pre_PostPCV$p, "BH", n = nrow(GPSC_VT_Pre_PostPCV))
names(GPSC_VT_Pre_PostPCV)[names(GPSC_VT_Pre_PostPCV) == "newcolumn"] <- "adjusted_p"
write.csv(GPSC_VT_Pre_PostPCV, file.path(base_path, "GPSC_VT_Pre_PostPCV.csv"), row.names = FALSE)

GPSC_NVT_Pre_PostPCV <- subset(GPSC_Pre_PostPCV, Vaccine_Status == "NVT")
GPSC_NVT_Pre_PostPCV$newcolumn <- p.adjust(GPSC_NVT_Pre_PostPCV$p, "BH", n = nrow(GPSC_NVT_Pre_PostPCV))
names(GPSC_NVT_Pre_PostPCV)[names(GPSC_NVT_Pre_PostPCV) == "newcolumn"] <- "adjusted_p"
write.csv(GPSC_NVT_Pre_PostPCV, file.path(base_path, "GPSC_NVT_Pre_PostPCV.csv"), row.names = FALSE)
```

------------------------------------------------------------------------

# Calculate AMR changes in proportion across three vaccine periods

## Install and load required packages

These are the packages required for analysis using this R script.

```{r install packages}
#install.packages("plyr")
#install.packages("pwr")

#Load packages
library(plyr)
library(pwr)
library(readr)
```

## Load input data

```{r load dataset}
# set base path for working directory
base_path <- getwd()

# set import file name
merged_data_file_name <- "merged_data.csv"

# set data file path
merged_data_file_path <- file.path(base_path, merged_data_file_name)

# load input data
merged_data <- read_csv(merged_data_file_path)

#Subset the data to include the information you need
AR_props <- subset(merged_data, select = c("Vaccine_Period", "Vaccine_Status", grep("PEN_Res.Meningital.$", colnames(merged_data), value = TRUE), grep("CHL_Res$", colnames(merged_data), value = TRUE), grep("COT_Res$", colnames(merged_data), value = TRUE), grep("ERY_Res$", colnames(merged_data), value = TRUE), grep("TET_Res$", colnames(merged_data), value = TRUE)))

#Convert back to a data frame
AR_props <- data.frame(AR_props)
# To analyse the data considering consider NVTs only or VTs (PCV serotypes) only, subset the data to include only NVTs and VTs.
#AR_props <- subset(AR_props, AR_props$Vaccine_Status %in% c("NVT"))
#AR_props <- subset(AR_props, AR_props$Vaccine_Status %in% c("PCV"))
```

## Create an output dataframe to store the results

```{r Output dataframe}
AR_Pre_PostPCV <- data.frame("Vaccine" = 0, "Vaccine_Period" = 0, "Antibiotic" = 0, "Pre-PCV_R" = 0, "Pre-PCV_total" = 0, "Post-PCV_R" = 0, "Post-PCV_total" = 0, "p" = 0)
Pre_R <- 0
Pre_S <- 0
Post_R <- 0
Post_S <- 0
Pre_R_proportion <- 0
Pre_S_proportion <- 0
Post_R_proportion <- 0
Post_S_proportion <- 0
non_susceptible <- c("I", "R")
```

## Define the vaccine periods

```{r Define the vaccine periods}
vaccine_period <- unique(AR_props$Vaccine_Period)
```

## To evaluate the changes in proportion count resistant and susceptible samples in the different vaccines periods to be compared. Then calculate the proportion. Perform chi-square test and calculate the power and significance.Adding 0.001 ensures that the denominator is not zero.

```{r Calculate the proportion of resistant isolates in the vaccine periods;Power test and Fishers exact test}
for (k in vaccine_period){
  AR_props_vaccine_period <- subset(AR_props, AR_props$Vaccine_Period != k)
  # calculate proportions of non-susceptible and susceptible samples for the vaccine periods
  for (j in grep("_Res", colnames(AR_props_vaccine_period), value = TRUE)) {
    Pre_Post_AR <- subset(AR_props_vaccine_period, select = c(j, "Vaccine_Period"))
    pcv <- unique(AR_props_vaccine_period$Vaccine_Period)
    
    Pre_R <- nrow(subset(Pre_Post_AR, Vaccine_Period == pcv[1][1] & Pre_Post_AR[, 1] %in% c("I", "R")))
    Pre_S <- nrow(subset(Pre_Post_AR, Vaccine_Period == pcv[1][1] & Pre_Post_AR[, 1] == "S"))
    Post_R <- nrow(subset(Pre_Post_AR, Vaccine_Period == pcv[2][1] & Pre_Post_AR[, 1] %in% c("I", "R")))
    Post_S <- nrow(subset(Pre_Post_AR, Vaccine_Period == pcv[2][1] & Pre_Post_AR[, 1] == "S"))
    
    Pre_R_proportion <- Pre_R / (Pre_R + Pre_S + 0.001) + 0.001
    Pre_S_proportion <- Pre_S / (Pre_R + Pre_S + 0.001) + 0.001
    Post_R_proportion <- Post_R / (Post_R + Post_S + 0.001) + 0.001
    Post_S_proportion <- Post_S / (Post_R + Post_S + 0.001) + 0.001
    
    AR <- matrix(c(Pre_R, Pre_S, Post_R, Post_S), nrow = 2, byrow = TRUE) # create matrix to store counts of non-susceptible and susceptible samples for the vaccine periods
    AR_proportion <- matrix(c(Pre_R_proportion, Pre_S_proportion, Post_R_proportion, Post_S_proportion), nrow = 2, byrow = TRUE) # create matrix to store the calculates proportions
    pwr <- pwr.chisq.test(w = ES.w2(AR_proportion), power = 0.8, df = 1, sig.level = 0.05) # calculate the power of the chi-square test with power and significance level
    
    # check the power (pwr) of the count of the resistant isolates before extracting your results.
    # check if the count is greater than the calculated power
    if (sum(AR) > pwr$N) {
      rownames(AR) <- pcv # assign the row names as the different vaccine periods
      colsum_AR <- colSums(AR) # total count of non-susceptible and susceptible samples in the vaccine periods
      row <- AR[1, , drop = FALSE] # extract the count of samples
      by2 <- matrix(c(row[1, 1], row[1, 2], colsum_AR[1] - row[1, 1], colsum_AR[2] - row[1, 2]), ncol = 2, byrow = TRUE) # contingency table for fisher's test
      p <- fisher.test(by2, alternative)$p.value # Fishers test on the contingency table by2
      pcv_compare <- unique(AR_props_vaccine_period$Vaccine_Period)
      AR_Pre_PostPCV <- rbind(AR_Pre_PostPCV, c(toString(pcv_compare), k, j, Pre_R, Pre_R + Pre_S, Post_R, Post_R + Post_S, p)) # matrix to store the results
    }
  }
}

AR_Pre_PostPCV  <-  AR_Pre_PostPCV [, -2]
```

## Write the results stratified by vaccine period to a csv file.

```{r Stratifying the results and write the output}
AR_Pre_PCV7 <- subset(AR_Pre_PostPCV, Vaccine == "Pre-PCV, PCV7")
AR_Pre_PCV7$newcolumn <- p.adjust(AR_Pre_PCV7$p, "BH", n = nrow(AR_Pre_PCV7))
names(AR_Pre_PCV7)[names(AR_Pre_PCV7) == "newcolumn"] <- "adjusted_p"
write.csv(AR_Pre_PCV7, file.path(base_path, "AMR_Pre_PCV7.csv"), row.names = FALSE)

AR_Pre_PCV13 <- subset(AR_Pre_PostPCV, Vaccine == "Pre-PCV, PCV13")
AR_Pre_PCV13$newcolumn <- p.adjust(AR_Pre_PCV13$p, "BH", n = nrow(AR_Pre_PCV13))
names(AR_Pre_PCV13)[names(AR_Pre_PCV13) == "newcolumn"] <- "adjusted_p"
write.csv(AR_Pre_PCV13, file.path(base_path, "AMR_Pre_PCV13.csv"), row.names = FALSE)

AR_PCV7_PCV13 <- subset(AR_Pre_PostPCV, Vaccine == "PCV7, PCV13")
AR_PCV7_PCV13$newcolumn <- p.adjust(AR_PCV7_PCV13$p, "BH", n = nrow(AR_PCV7_PCV13))
names(AR_PCV7_PCV13)[names(AR_PCV7_PCV13) == "newcolumn"] <- "adjusted_p"
write.csv(AR_PCV7_PCV13, file.path(base_path, "AMR_PCV7_PCV13.csv"), row.names = FALSE)
```
