---
title: "Create Plot"
output: html_document
date: "2024-03-11"
purpose: Generating the prevalence of GPSC and Serotype in different vaccination periods, and AMR profile for each GPSC
---

------------------------------------------------------------------------

# Generating the prevalence of GPSC and serotype in different vaccination periods

## Install and load required packages

These are the packages required for the R script session.

```{r install_packages_1}
#install.packages("ggplot2")
#install.packages("ggpattern")

library(ggplot2)
library(ggpattern)
```

## Functions

This function is to generate a frequency table based on **GPSC**, **Serotype** and **Vaccine_Period**.

```{r load_functions_1}
Generate_GPSC_Serotype_Vaccine_Freqtable <- function(metadata_df,
                                                     pair_colname1 = "GPSC", pair_colname2 = "In_Silico_Serotype",
                                                     pair_colname3 = "Vaccine_Period",
                                                     n_output_col = 6,
                                                     output_colnames = c("GPSC", "Serotype", "Vaccine_Period", "Count", "Proportion", "Pair_Name")) {
  for (isolate in 1:nrow(metadata_df)) {
    metadata_df[isolate, "pair"] <- paste(metadata_df[isolate, pair_colname1],
                                         metadata_df[isolate, pair_colname2],
                                         metadata_df[isolate, pair_colname3], sep = "_")
  }
  pair_freq <- as.data.frame(table(metadata_df$pair))
  colnames(pair_freq) <- c("name", "freq")
  
  dsize <- matrix(0, nrow = nrow(pair_freq), ncol = n_output_col)
  gpsc_sero_vacperiod_df <- as.data.frame(dsize)
  colnames(gpsc_sero_vacperiod_df) <- output_colnames
  
  for (i in 1:nrow(pair_freq)) {
    gpsc <- strsplit(as.character(pair_freq[i, 1]), "_")[[1]][1]
    pair_freq$gpsc[i] <- gpsc
    sero <- strsplit(as.character(pair_freq[i, 1]), "_")[[1]][2]
    vac_period <- strsplit(as.character(pair_freq[i, 1]), "_")[[1]][3]
    gpsc_sero_vacperiod_df[i, "GPSC"] <- gpsc
    gpsc_sero_vacperiod_df[i, "Serotype"] <- sero
    gpsc_sero_vacperiod_df[i, "Vaccine_Period"] <- vac_period
    gpsc_sero_vacperiod_df[i, "Pair_Name"] <- as.character(pair_freq[i, "name"])
    
  }
  uniq_pair <- unique(metadata_df$pair)
  ### each row is a population-serotype-gpsc pair
  num_pre_vac <- length(which(metadata_df[, pair_colname3] == "Pre-PCV"))
  num_PCV7 <- length(which(metadata_df[, pair_colname3] == "PCV7"))
  num_PCV13 <- length(which(metadata_df[, pair_colname3] == "PCV13"))
  
  for (p in uniq_pair){
    selected_pair <- metadata_df[which(metadata_df$pair == p), ]
    t_count <- nrow(selected_pair)
    gpsc_sero_vacperiod_df[which(gpsc_sero_vacperiod_df$Pair_Name == p), "Count"] <- t_count
    if (gpsc_sero_vacperiod_df[which(gpsc_sero_vacperiod_df$Pair_Name == p), "Vaccine_Period"] == "pre-PCV") {
      gpsc_sero_vacperiod_df[which(gpsc_sero_vacperiod_df$Pair_Name == p), "Proportion"] <- t_count / num_pre_vac
    } else if (gpsc_sero_vacperiod_df[which(gpsc_sero_vacperiod_df$Pair_Name == p), "Vaccine_Period"] == "PCV7") {
      gpsc_sero_vacperiod_df[which(gpsc_sero_vacperiod_df$Pair_Name == p), "Proportion"] <- t_count / num_PCV7
    } else {
      gpsc_sero_vacperiod_df[which(gpsc_sero_vacperiod_df$Pair_Name == p), "Proportion"] <- t_count / num_PCV13
    }
  }
  
  return(gpsc_sero_vacperiod_df)
}
```

## Load input data

The `serotype__colours_patterns.csv` file include the information about the colour coding for each serotype as well as the vaccine type information.\
Note that the vaccine type information is based on PCV13 implemented in Argentina, which might need some adjustment for your studied population.

```{r load_dataset_1}
# set base path for working directory
base_path <- getwd()

# set import file names
merged_data_file_name <- "merged_data.csv"
serotype_colorcode_file_name <- "serotype__colours_patterns.csv"

# set data file paths
merged_data_file_path <- file.path(base_path, merged_data_file_name)
serotype_colorcode_file_path <- file.path(base_path, serotype_colorcode_file_name)

# load input data
merged_data <- read.csv(merged_data_file_path)
colnames(merged_data)[colnames(merged_data) == "Serotype"] <- "In_Silico_Serotype" # rename column Serotype to In_Silico_Serotype for the script
data_frame <- as.data.frame(merged_data)
serotype_colorcode <- read.table(serotype_colorcode_file_path,
                                 sep = ",", comment.char = "&", header = TRUE)
```

## Check if terminology is consistent

```{r check_terms_1}
data_frame[which(data_frame$GPSC == "_"), "GPSC"] <- "Others"
data_frame[which(data_frame$GPSC == "904;9"), "GPSC"] <- "9"
data_frame[which(data_frame$In_Silico_Serotype == "_"), "In_Silico_Serotype"] <- "NT"
## Might need to check Vaccine Period
```

## Transform into frequency table

In the output table, each row is a unique combination of GPSC-Serotype-Vaccine_Period. Note that the column names for GPSC, Serotype, and Vaccine_Period should be:

-   GPSC: "GPSC", otherwise please specify the column names by *pair_colname1="Your_GPSC_Column_Name"*\
-   Serotype: "In_Silico_Serotype", otherwise please specify the column names by *pair_colname2="Your_Serotype_Column_Name"*\
-   Vaccine_Period: "Vaccine_Period", otherwise please specify the column names by *pair_colname2="Your_Vaccine_Period_Column_Name"*

```{r generate_pair_frequency_table_1}
## Need to generate/transform the dataframe to each row as a GPSC-Serotype-Vaccine_Period pair with corresponding count/proportion
GPS_freq_table <- Generate_GPSC_Serotype_Vaccine_Freqtable(metadata_df = data_frame,
                                                           pair_colname1 = "GPSC",
                                                           pair_colname2 = "In_Silico_Serotype",
                                                           pair_colname3 = "Vaccine_Period")

# Add serotype colour coding information to the frequency table
GPS_freq_table_sero_code_merge_gpsc <- merge(GPS_freq_table, serotype_colorcode, by = "Serotype")

# Merge rare GPSC to "Others"
target_gpsc <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "12", "13", "15", "16", "18", "23", "26", "31",
                 "32", "37", "44", "47", "50", "61", "70", "95", "113", "386", "657")
GPS_freq_table_sero_code_merge_gpsc[which(!GPS_freq_table_sero_code_merge_gpsc$GPSC %in% target_gpsc), "GPSC"] <- "Others"

head(GPS_freq_table_sero_code_merge_gpsc)
```

## Setting the axis order for the plot

```{r set_plot_label_order_1}
### Set GPSC order
GPSC_order <- c(target_gpsc, "Others")
GPS_freq_table_sero_code_merge_gpsc$GPSC <- factor(GPS_freq_table_sero_code_merge_gpsc$GPSC, levels = GPSC_order)

### Set vaccine period order
vaccine_period_order <- c("Pre-PCV", "PCV7", "PCV13")
GPS_freq_table_sero_code_merge_gpsc$Vaccine_Period <- factor(GPS_freq_table_sero_code_merge_gpsc$Vaccine_Period, levels = vaccine_period_order)

### Set serotype order by vaccine type
# merge 06A and 6E(6A)
GPS_freq_table_sero_code_merge_gpsc[which(GPS_freq_table_sero_code_merge_gpsc$Serotype == "6E(6A)"), "Serotype"] <- "06A"
GPS_freq_table_sero_code_merge_gpsc[which(GPS_freq_table_sero_code_merge_gpsc$Serotype == "6E(6B)"), "Serotype"] <- "06B" 
GPS_freq_table_sero_code_merge_gpsc[which(GPS_freq_table_sero_code_merge_gpsc$Serotype == "23B1"), "Serotype"] <- "23B" 


PCV7_serotype <- c("4", "6E(6B)", "09V", "14", "18C", "19F", "23F")
PCV13_serotype <- c("1", "3", "5", "06A", "7F", "19A")
PCV_serotype <- c(PCV7_serotype, PCV13_serotype)
all_serotype <- unique(GPS_freq_table_sero_code_merge_gpsc$Serotype)
other_serotype <- all_serotype[!(all_serotype %in% PCV_serotype)]

serotype_order <- c(PCV7_serotype, PCV13_serotype, other_serotype) # add an blank element to separate VT and NVT
GPS_freq_table_sero_code_merge_gpsc$Serotype <- factor(GPS_freq_table_sero_code_merge_gpsc$Serotype, levels = serotype_order)

### Set serotype colour order
filled_colours <- GPS_freq_table_sero_code_merge_gpsc$Serotype__colour
names(filled_colours) <- GPS_freq_table_sero_code_merge_gpsc$Serotype

### Set serotype pattern order
filled_patterns <- GPS_freq_table_sero_code_merge_gpsc$ggpattern
names(filled_patterns) <- GPS_freq_table_sero_code_merge_gpsc$Vtype
```

## Generate GPSC-Serotype-Vaccine_Period proportion plot

```{r plot_1}
GPS_freq_table_plot <- ggplot(GPS_freq_table_sero_code_merge_gpsc, aes(x = Vaccine_Period, y = Count, fill = Serotype, pattern = Vtype)) + #fill=Serotype, pattern=Serotype
  geom_bar_pattern(position = "stack", stat = "identity", pattern_density = 0.2, pattern_spacing = 0.2, pattern_key_scale_factor = 1.2, pattern_size = 0) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 20),
        axis.text.y = element_text(size = 25),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25),
        plot.caption = element_text(size = 25)) +
  theme(strip.placement = "outside",
        strip.background = element_rect(fill = NA, color = "white"),
        strip.text = element_text(size = 25),
        panel.spacing = unit(-.01, "cm"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.key.size = unit(0.8, "cm")) +
  xlab("GPSC") +
  #ylim(c(0, 0.3)) +
  facet_grid(~ GPSC, switch = "x") +
  scale_fill_manual(values = filled_colours) +
  scale_pattern_manual(values = filled_patterns, guide = "none") +
  labs(caption = "*Non-PCV13 Type are labeled with stripes")

#GPS_freq_table_plot
```

## Save plot to .png file

```{r save_plot_1}
output_path <- file.path(base_path, "GPS_freq_table_plot.png")
png(output_path, width = 2500, height = 1000)
GPS_freq_table_plot
dev.off()
```

------------------------------------------------------------------------

# AMR profile for each GPSC

## Install and load required packages

These are the packages required for the R script session.

```{r install_packages_2}
#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("gridExtra")
#install.packages("scales")
#install.packages("cowplot")

library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(scales)
library(cowplot)
```

## Functions

This function is to generate a frequency table based on **GPSC**, **Serotype** and, **Vaccine_Period**.

```{r load_functions_2}
percent_amr <- function(amr_df, amr, ris = c("I", "R")) {
  num_pen_men_r <- length(which(amr_df[, amr] %in% ris))
  pen_men_percent <- num_pen_men_r / nrow(amr_df)
  return(pen_men_percent)
}
```

## Load input data

```{r load_dataset_2}
# set base path for working directory
base_path <- getwd()

# set import file name
merged_data_file_name <- "merged_data.csv"

# set data file path
merged_data_file_path <- file.path(base_path, merged_data_file_name)

# Load input data
merged_data <- read.csv(merged_data_file_path)
data_frame <- as.data.frame(merged_data)

# Select necessary columns: GPSC, PEN_Res(Meningital), CHL_Res, TET_Res, COT_Res, mefA, ermB
AMRs <- c("PEN_Res.Meningital.", "CHL_Res", "TET_Res", "COT_Res", "ERY_Res")
AMR_data_frame <- data_frame[, c("GPSC", AMRs)]

# Merge other GPSCs
target_gpsc <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "12", "13", "15", "16", "18", "23", "26", "31",
                 "32", "37", "44", "47", "50", "61", "70", "95", "113", "386", "657")

AMR_data_frame_merge_gpsc <- AMR_data_frame
AMR_data_frame_merge_gpsc[which(!AMR_data_frame_merge_gpsc$GPSC %in% target_gpsc), "GPSC"] <- "Others"
```

## Generate GPSC-AMR_percentage table

```{r AMR_percentage_table_2}
output_colnames <- c("GPSC", "PEN_MEN", "CHL", "TET", "COT", "ERY")
unique_gpsc <- unique(AMR_data_frame_merge_gpsc$GPSC)
dsize <- matrix(0, nrow = length(unique_gpsc), ncol = ncol(AMR_data_frame))
AMR_percentage_df <- as.data.frame(dsize)
colnames(AMR_percentage_df) <- output_colnames

IR <- c("I", "R")

for (i in 1:length(unique_gpsc)){
  gpsci <- unique_gpsc[i]
  gpsc_df <- AMR_data_frame_merge_gpsc[which(AMR_data_frame_merge_gpsc$GPSC == gpsci), ]
  #PEN_MEN
  percent_PEN_MEN <- percent_amr(amr_df = gpsc_df, amr = AMRs[1], ris = IR)
  #CHL
  percent_CHL <- percent_amr(amr_df = gpsc_df, amr = AMRs[2], ris = IR)
  #TET
  percent_TET <- percent_amr(amr_df = gpsc_df, amr = AMRs[3], ris = IR)
  #COT
  percent_COT <- percent_amr(amr_df = gpsc_df, amr = AMRs[4], ris = IR)
  #ermB
  percent_ERY <- percent_amr(amr_df = gpsc_df, amr = AMRs[5], ris = IR)

  AMR_percentage_df[i, "GPSC"] <- gpsci
  AMR_percentage_df[i, "PEN_MEN"] <- percent_PEN_MEN
  AMR_percentage_df[i, "CHL"] <- percent_CHL
  AMR_percentage_df[i, "TET"] <- percent_TET
  AMR_percentage_df[i, "COT"] <- percent_COT
  AMR_percentage_df[i, "ERY"] <- percent_ERY
}

# Transform AMR table to long format which is a requirement for ggplot
AMR_percentage_df_long <- AMR_percentage_df %>% gather(AMR, Percentage, PEN_MEN:ERY)
```

## Setting the order of the plot

```{r set_order_2}
# Set AMR order
AMR_order <- c("TET", "COT", "ERY", "CHL", "PEN_MEN")
AMR_percentage_df_long$AMR <- factor(AMR_percentage_df_long$AMR, levels = AMR_order)

# Set GPSC order
gpscs <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "12", "13", "15", "16", "18", "23", "26", "31",
           "32", "37", "44", "47", "50", "61", "70", "95", "113", "386", "657", "Others")
```

## Plot GPSC1 AMR profile, which is the leftmost plot

This plot will have both x and y-axis label

```{r GPSC1_AMR_plot_2}
GPSC1_amr <- AMR_percentage_df_long[which(AMR_percentage_df_long$GPSC == "1"), ]
left_amr_ggplot <-
  ggplot(GPSC1_amr, aes(x = AMR, y = Percentage)) +
  geom_bar(stat = "identity", fill = "#f68060", alpha = .6, width = .4) +
  scale_y_continuous(
    limits = c(0, 1.2),
    breaks = seq(0, 1, by = 0.5)
  ) +
  coord_flip() +
  xlab("") +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  ggtitle("GPSC1") +
  theme(
    axis.text.y = element_text(size = 50),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 50),
    plot.title = element_text(size = 50)
  )
```

## Plot all other GPSCs AMR profile

```{r AMR_plots_2}
for (i in 1:length(gpscs)){
  gpsc_variable <- paste0("GPSC_", gpscs[i], "_amr")
  assign(gpsc_variable, AMR_percentage_df_long[which(AMR_percentage_df_long$GPSC == gpscs[i]), ])
  
  gpsc_amr_plot_variable <- paste0(gpsc_variable, "_plot")
  gpsc_title <- paste0("GPSC", gpscs[i])
  new_gpsc_amr_plot <-
    ggplot(get(gpsc_variable), aes(x = AMR, y = Percentage)) +
    geom_bar(stat = "identity", fill = "#f68060", alpha = .6, width = .4) +
    scale_y_continuous(
      limits = c(0, 1.2),
      breaks = seq(0, 1, by = 0.5),
      labels = label_number(accuracy = 0.1)
    ) +
    coord_flip() +
    labs(x = NULL, y = NULL) +
    ggtitle(gpsc_title) +
    theme_bw() +
    theme(
      axis.text.y = element_blank(),
      panel.background = element_rect(fill = "white"),
      panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      axis.text.x = element_text(size = 50),
      plot.title = element_text(size = 50)
    )
  
  assign(gpsc_amr_plot_variable, new_gpsc_amr_plot)
}

```

## Bind all the AMR profile together

```{r bind_plots_2}
wid <- rep(0.8 / 30, 30)
output_path <- file.path(base_path, "AMR_plot.png")
png(output_path, width = 9500, height = 500)
plot_grid(left_amr_ggplot, GPSC_2_amr_plot, GPSC_3_amr_plot,
          GPSC_4_amr_plot, GPSC_5_amr_plot, GPSC_6_amr_plot,
          GPSC_7_amr_plot, GPSC_8_amr_plot, GPSC_9_amr_plot,
          GPSC_10_amr_plot, GPSC_12_amr_plot, GPSC_13_amr_plot,
          GPSC_15_amr_plot, GPSC_16_amr_plot, GPSC_18_amr_plot,
          GPSC_23_amr_plot, GPSC_26_amr_plot, GPSC_31_amr_plot,
          GPSC_32_amr_plot, GPSC_37_amr_plot, GPSC_44_amr_plot,
          GPSC_47_amr_plot, GPSC_50_amr_plot, GPSC_61_amr_plot,
          GPSC_70_amr_plot, GPSC_95_amr_plot, GPSC_113_amr_plot,
          GPSC_386_amr_plot, GPSC_657_amr_plot, GPSC_Others_amr_plot,
          align = "h", ncol = 30, rel_widths = c(1.5 / 30, wid))
dev.off()
```
